### 2020-07-01 notebooks/old_notebooks/us_explore2.Rmd

---
title: "Intersection COVID-19: Explore data"
date: "`r Sys.time()`"
categories:
  - covid-19
tags: 
  - covid-19
  - Italy
lastmod: "`r Sys.time()`"all
keywords: []
description: ''
output:
  html_document:
    df_print: kable
    toc: true
    theme: journal
---

# Objective

```{r}
library(facetscales)
library(covid19census)
library(tidyverse)
library(ggcorrplot)
library(GGally)
library(gganimate)
library(ggcorrplot)
library(broom)
library(purrr)
library(knitr)
library(kableExtra)
library(htmlwidgets)
library(plotly)
library(stringr)
library(ggridges)
library(xtable)
theme_set(theme_bw())
```

# Italy Table

## flu block
```{r}
data_it <- getit_all()

today_it <- 
  data_it %>% 
    filter(date == max(date))

new_order_it <- c("n", "deaths", "pop_km2", "lat", "chronic_diab", "chronic_bronc", "chronic_hyp")

tab_it_flu <- 
  today_it %>% 
  select(region, deaths, lat, imm65, pop_km2, chronic_diab, chronic_bronc, chronic_hyp, pop_tot) %>% 
  mutate_if(is.numeric, round, 1) %>% 
  # 53 is median value
  mutate(imm65= if_else(imm65 > 50, "high53", "low53")) %>% 
  mutate( deaths = deaths/pop_tot * 100000) %>% 
  group_by(imm65) %>% 
  mutate(n = n()) %>% 
  summarise_if(is.numeric, funs(mean, sd)) %>% 
  pivot_longer(-imm65) %>% 
  separate(name, c("var", "metric"), "_(?=(mean)|(sd))") %>% 
  unite(imm65, imm65, metric ) %>% 
  pivot_wider(names_from = imm65, values_from = value) %>% 
  mutate_if(is.numeric, round,1)  %>% 
  mutate(
     "<53% (St.Dev.)"= paste(low53_mean,"(", low53_sd, ")"),
     ">53% (St.Dev.)"= paste(high53_mean,"(",high53_sd, ")")) %>% 
  select(var, "<53% (St.Dev.)", ">53% (St.Dev.)") %>% 
  filter(var != "pop_tot") %>%
  arrange(match(var, !!new_order_it)) %>% 
  mutate(var = recode(var,
    "n" = "Number of counties",
    "deaths" = "COVID-19 death rate (per 100,000 people)",
    "lat" = "Latitude",
    "chronic_diab" = "% chronic diabetes",
    "chronic_bronc" =  "% chronic bronchitis",
    "chronic_hyp" = "% chronic hypertension",
    "pop_km2" = "Population density (people per squared km)"
    ))

```

### tot

```{r}
new_order_it <- c("n", "deaths", "pop_km2", "lat", "chronic_diab", "chronic_bronc", "chronic_hyp")

tab_it_tot <- 
  today_it %>% 
  select(region, deaths, lat, imm65, pop_km2, chronic_diab, chronic_bronc, chronic_hyp, pop_tot) %>% 
  mutate_if(is.numeric, round, 1) %>% 
  # 53 is median value
  mutate( deaths = deaths/pop_tot * 100000,
          all_d = "all") %>% 
  group_by(all_d) %>% 
  mutate(n = n()) %>% 
  summarise_if(is.numeric, funs(mean, sd)) %>% 
  pivot_longer(-all_d) %>% 
  separate(name, c("var", "metric"), "_(?=(mean)|(sd))") %>% 
  unite(all_d, all_d, metric ) %>% 
  pivot_wider(names_from = all_d, values_from = value) %>% 
  mutate_if(is.numeric, round, 1)  %>%

  mutate(
     "All (St.Dev.)" = paste(all_mean,"(", all_sd, ")")) %>% 
  select(var, "All (St.Dev.)") %>% 
  filter(!var %in% c("pop_tot", "imm65"))%>%
  arrange(match(var, !!new_order_it)) %>% 
  mutate(var = recode(var,
    "n" = "Number of counties",
    "deaths" = "COVID-19 death rate (per 100,000 people)",
    "lat" = "Latitude",
    "chronic_diab" = "% chronic diabetes",
    "chronic_bronc" =  "% chronic bronchitis",
    "chronic_hyp" = "% chronic hypertension",
    "pop_km2" = "Population density (people per squared km)"
    ))

tab_it_all <- inner_join(tab_it_tot, tab_it_flu, by = "var")

print(
  xtable(
    tab_it_all
  ),
  include.rownames=FALSE,
  comment = FALSE,
  align = rep("c", ncol(tab_it))
)

```

# US Table

```{r warning=FALSE}

logit = function(p){ log(p) - log(1-p) }

# Thi are the variable that are initially selected from the dataframe
# var new to select -------
to_select<- c("date", "county", "state", "fips", "cases", "deaths", "perc_families", 
"perc_family_only_onep", "perc_edu_bachelor", 
"perc_withinternet", "total_pop", "perc_imm65", "total_imm65", 
"total_beds", "ratio_beds", "urban",  
"perc_alzheimer_dementia", "perc_asthma", "perc_atrial_fibrillation", 
"perc_cancer_breast", "perc_cancer_colorectal", "perc_cancer_lung", 
"perc_ch_obstructive_pulm", "perc_chronic_kidney_disease", "perc_depression", 
"perc_diabetes", "perc_heart_failure", "perc_hypertension", "perc_ischemic_heart_disease", "perc_obesity", "perc_osteoporosis", "perc_rheumatoid_arthritis", 
 "perc_stroke", "perc_tobacco_use", 
"median_income", "pm2.5", "summer_temp", "summer_hum", "winter_temp", 
"winter_hum", "age65_over", "median_age", "sex_ratio", "child_dependency", 
"perc_black", "perc_lat", "perc_white", "perc_asian", "perc_island", 
"perc_other", "perc_two_more_races", "total_tests", "days_f0"
)

# df1   -----------------------------
df1 <- getus_all()

# df2 ------------------------------
df2 <-
  df1 %>%
  # calculate age65_over and perc_black, white and latinos, total tests
  mutate(age65_over = rowSums(dplyr::select(., matches("perc_[6][5-9]|[7|8][0-9]")))) %>%
  mutate(urban = if_else(urban == "Urban", 1, 0)) %>%
  mutate(perc_black = total_black / total_pop * 100) %>%
  mutate(perc_white = total_white / total_pop * 100) %>%
  mutate(perc_lat = total_latino / total_pop * 100) %>%
  mutate(total_tests = positive + negative) %>%

  # calculate day since first case
  mutate(f_date = case_when(cases >= 1 ~ date)) %>%
  group_by(fips) %>%
  mutate(f_date = min(f_date, na.rm = TRUE), days_f0 = as.numeric(date - f_date)) %>%
  ungroup() %>%
  mutate(days_f0 = if_else(is.finite(days_f0), days_f0, NA_real_)) %>%

  # newly calculated-----------------------------------  
  # total beds in ratio
  mutate(ratio_beds = total_beds/total_pop) %>% 
  # asian are here
  mutate(perc_asian = total_asian / total_pop * 100) %>%
  mutate(perc_island = total_pacific_islander / total_pop * 100) %>%
  mutate(perc_native = total_native / total_pop * 100) %>%
  mutate(perc_other = total_other_race / total_pop * 100) %>%
  mutate(perc_two_more_races = total_two_more_races / total_pop * 100) %>%

  # family with one parent together
  mutate(perc_family_only_onep = perc_families_only_female + perc_families_only_male) %>% 
  mutate(total_imm65 = age65_over/100 *perc_imm65) %>%

  # select variables ----------
  dplyr::select(!!to_select) %>%

  # perc divided by 100
  mutate_at(vars(starts_with("perc")), function(x) x / 100) %>%
  mutate(age65_over = age65_over / 100) %>%

  # we remove NAs
  na.omit() %>%

  # at least 1 death
  filter(deaths >= 1)

# all_dat -------------
all_dat <- df2 %>%
  # make variables numeric
  mutate_at(vars(
    -date, -county, -state, -fips), 
     as.numeric
  ) %>%
  # rename
  # add XX_ to all the variables excepts those
  rename_at(
    vars(
      -date, -county, -state, -fips, -cases, -deaths,
      -perc_imm65, -cases, -total_pop, -total_imm65
    ),
    ~ paste0("XX_", .)
  ) %>%
  rename_at(vars(perc_imm65, total_imm65), ~ paste0("ZZ_", .)) %>%
  rename_at(vars(deaths), ~ paste0("YY_", .)) %>%
  rename_at(vars(cases), ~ paste0("NC_", .)) %>%
  rename_at(vars(total_pop), ~ paste0("NP_", .)) %>%
  # calculate logitZZ
  mutate(logitZZ_perc_imm65 = logit(ZZ_perc_imm65)) 
  # mutate(logitZZ_total_imm65 = log(ZZ_total_imm65) - log(1 - ZZ_total_imm65))

# today_dat ------------------------------
today_dat <- all_dat %>%
  filter(date == max(date))
```

```{r}
new_order <- c(
  "n",
  "death_rate",
  "XX_perc_families",
  "XX_perc_family_only_onep",
  "XX_perc_edu_bachelor",
  "XX_perc_withinternet",
  "XX_ratio_beds",
  "XX_urban",
  "XX_perc_alzheimer_dementia",
  "XX_perc_asthma",
  "XX_perc_atrial_fibrillation",
  "XX_perc_cancer_breast",
  "XX_perc_cancer_colorectal",
  "XX_perc_cancer_lung",
  "XX_perc_ch_obstructive_pulm",
  "XX_perc_chronic_kidney_disease",
  "XX_perc_depression",
  "XX_perc_diabetes",
  "XX_perc_heart_failure",
  "XX_perc_hypertension",
  "XX_perc_ischemic_heart_disease",
  "XX_perc_obesity",
  "XX_perc_osteoporosis",
  "XX_perc_rheumatoid_arthritis",
  "XX_perc_stroke",
  "XX_perc_tobacco_use",
  "XX_median_income",
  "XX_pm2.5",
  "XX_summer_temp",
  "XX_summer_hum",
  "XX_winter_temp",
  "XX_winter_hum",
  "XX_median_age",
  "XX_age65_over",
  "XX_sex_ratio",
  "XX_child_dependency",
  "XX_perc_black",
  "XX_perc_lat",
  "XX_perc_white",
  "XX_perc_asian",
  "XX_perc_island",
  "XX_perc_other",
  "XX_perc_two_more_races",
  "XX_days_f0"
)
```

## flu blocks

```{r}

# RATIO BEDS X THOUSANDS
pm2.5_name <- expression("% average PM2.5(" ~ mu ~ g / m^3 ~ ")")

tab_us_flu <- today_dat %>% 
   mutate(median = median(ZZ_perc_imm65)) %>% 
   select(matches("ZZ|XX|YY|NP")) %>% 
   select(
     -ZZ_total_imm65,
     -logitZZ_perc_imm65,
     -XX_total_tests,
     -XX_total_tests,
     -XX_total_beds

     ) %>% 
   mutate_at(vars(matches("perc")), ~ . * 100) %>% 
   mutate(ZZ_perc_imm65 = if_else(ZZ_perc_imm65 > 46, "high46", "low46")) %>%

   mutate(
     death_rate = YY_deaths / NP_total_pop * 100000,
     XX_urban = XX_urban * 100,
     XX_ratio_beds = XX_ratio_beds * 1000,
     XX_median_income	= XX_median_income / 1000,
     XX_sex_ratio = XX_sex_ratio * 100,
     XX_child_dependency = XX_sex_ratio * 100,
     ) %>%

   group_by(ZZ_perc_imm65) %>% 
   mutate(n = n()) %>% 
   summarize_if(is.numeric, funs(mean, sd)) %>%

   pivot_longer(-ZZ_perc_imm65) %>% 
   separate(name, c("var", "metric"), "_(?=(mean)|(sd))") %>% 
   unite(imm65, ZZ_perc_imm65, metric ) %>%

   pivot_wider(names_from = imm65, values_from = value) %>% 
   mutate_if(is.numeric, round,1)  %>% 
   mutate(
     "<46% (St.Dev.)"= paste(low46_mean,"(", low46_sd, ")"),
     ">46% (St.Dev.)"= paste(high46_mean,"(",high46_sd, ")")) %>% 
  select(var, "<46% (St.Dev.)", ">46% (St.Dev.)") %>%
  filter(!var %in% c("NP_total_pop", "YY_deaths")) %>% 
  arrange(match(var, !!new_order)) %>% 
  mutate(var = recode(var,
    "n" = "Number of counties",
    "death_rate" = "COVID-19 death rate (per 100,000 people)",
    "XX_perc_families" = "% families",
    "XX_perc_family_only_onep" = "% families with only one parent",
    "XX_perc_edu_bachelor" =   "% with bachelor",
    "XX_perc_withinternet" = "% with Internet",
    "XX_ratio_beds" = "Rate of hospital beds (per 100,000 people)",
    "XX_urban" = "% urban counties",
    "XX_perc_alzheimer_dementia" = "% with Alzheimer's disease",
    "XX_perc_asthma" = "% with asthma",
    "XX_perc_atrial_fibrillation" = "% with atrial fibrillation",
    "XX_perc_cancer_breast" = "% with breast cancer",
    "XX_perc_cancer_colorectal" = "% with colorectal cancer",
    "XX_perc_cancer_lung" = "% with lung cancer",
    "XX_perc_ch_obstructive_pulm" = "% with obstructive pulmonary disease",
    "XX_perc_chronic_kidney_disease" = "% with chronic kidney disease",
    "XX_perc_depression" = "% with depression",
    "XX_perc_diabetes" = "% with diabetes",
    "XX_perc_heart_failure" = "% with heart failure",
    "XX_perc_hypertension" = "% with hypertension",
    "XX_perc_ischemic_heart_disease" = "% with ischemic heart disease",
    "XX_perc_obesity" = "% with obesity",
    "XX_perc_osteoporosis" = "% with osteoporosis",
    "XX_perc_rheumatoid_arthritis" = "% with rheumatoid arthritis", 
    "XX_perc_stroke" = "% stroke transient ischemic attack",
    "XX_perc_tobacco_use" =  "% using tobacco",
    "XX_median_income" = "% median income ($1,000)",
    "XX_pm2.5" = "% average PM2.5(microg/m3)",
    "XX_summer_temp" = "Summer temperature (K)",
    "XX_summer_hum" =  "Summer humidity (mixing ratio)", 
    "XX_winter_temp" = "Winter temperature (K)",
    "XX_winter_hum" = "Winter humidity (mixing ratio)",
    "XX_median_age" = "% median age",
    "XX_age65_over" = "Number 65 years of age",
    "XX_sex_ratio" = "Sex ratio",
    "XX_child_dependency" = "Child dependency ratio",
    "XX_perc_black" =   "% black",
    "XX_perc_lat" =  "% latino",
    "XX_perc_white" = "% white",
    "XX_perc_asian" = "% asian",
    "XX_perc_island" = "% island native",
    "XX_perc_other" =   "% other",
    "XX_perc_two_more_races" = "% two more races",
    "XX_days_f0" =  "Number of days since first case"

  )
  )

# total test need to be added after
```

## tot

```{r US_ALL}
tab_us_tot <-
  today_dat %>% 
   mutate(median = median(ZZ_perc_imm65)) %>% 
   select(matches("ZZ|XX|YY|NP")) %>% 
   select(
     -ZZ_total_imm65,
     -logitZZ_perc_imm65,
     -XX_total_tests,
     -XX_total_tests,
     -XX_total_beds

     ) %>% 
   mutate_at(vars(matches("perc")), ~ . * 100) %>% 
   mutate(ZZ_perc_imm65 = if_else(ZZ_perc_imm65 > 46, "high46", "low46")) %>%

   mutate(
     death_rate = YY_deaths / NP_total_pop * 100000,
     XX_urban = XX_urban * 100,
     XX_ratio_beds = XX_ratio_beds * 1000,
     XX_median_income	= XX_median_income / 1000,
     XX_sex_ratio = XX_sex_ratio * 100,
     XX_child_dependency = XX_sex_ratio * 100,
     ) %>%

   group_by(all_d) %>% 
   mutate(n = n()) %>% 
   summarize_if(is.numeric, funs(mean, sd)) %>%

   pivot_longer(-all_d) %>% 
   separate(name, c("var", "metric"), "_(?=(mean)|(sd))") %>% 
   unite(all_d, all_d, metric ) %>% 
   pivot_wider(names_from = all_d, values_from = value) %>% 
    mutate_if(is.numeric, round, 1)  %>%
  mutate(
     "All (St.Dev.)" = paste(all_d_mean,"(", all_d_sd, ")")) %>% 
  select(var, "All (St.Dev.)") %>% 
  mutate_if(is.numeric, round, 1)  %>% 
  filter(!var %in% c("NP_total_pop", "YY_deaths")) %>% 
  arrange(match(var, !!new_order)) %>% 
 mutate(var = recode(var,
    "n" = "Number of counties",
    "death_rate" = "COVID-19 death rate (per 100,000 people)",
    "XX_perc_families" = "% families",
    "XX_perc_family_only_onep" = "% families with only one parent",
    "XX_perc_edu_bachelor" =   "% with bachelor",
    "XX_perc_withinternet" = "% with Internet",
    "XX_ratio_beds" = "Rate of hospital beds (per 100,000 people)",
    "XX_urban" = "% urban counties",
    "XX_perc_alzheimer_dementia" = "% with Alzheimer's disease",
    "XX_perc_asthma" = "% with asthma",
    "XX_perc_atrial_fibrillation" = "% with atrial fibrillation",
    "XX_perc_cancer_breast" = "% with breast cancer",
    "XX_perc_cancer_colorectal" = "% with colorectal cancer",
    "XX_perc_cancer_lung" = "% with lung cancer",
    "XX_perc_ch_obstructive_pulm" = "% with obstructive pulmonary disease",
    "XX_perc_chronic_kidney_disease" = "% with chronic kidney disease",
    "XX_perc_depression" = "% with depression",
    "XX_perc_diabetes" = "% with diabetes",
    "XX_perc_heart_failure" = "% with heart failure",
    "XX_perc_hypertension" = "% with hypertension",
    "XX_perc_ischemic_heart_disease" = "% with ischemic heart disease",
    "XX_perc_obesity" = "% with obesity",
    "XX_perc_osteoporosis" = "% with osteoporosis",
    "XX_perc_rheumatoid_arthritis" = "% with rheumatoid arthritis", 
    "XX_perc_stroke" = "% stroke transient ischemic attack",
    "XX_perc_tobacco_use" =  "% using tobacco",
    "XX_median_income" = "% median income ($1,000)",
    "XX_pm2.5" = "% average PM2.5(microg/m3)",
    "XX_summer_temp" = "Summer temperature (K)",
    "XX_summer_hum" =  "Summer humidity (mixing ratio)", 
    "XX_winter_temp" = "Winter temperature (K)",
    "XX_winter_hum" = "Winter humidity (mixing ratio)",
    "XX_median_age" = "% median age",
    "XX_age65_over" = "Number 65 years of age",
    "XX_sex_ratio" = "Sex ratio",
    "XX_child_dependency" = "Child dependency ratio",
    "XX_perc_black" =   "% black",
    "XX_perc_lat" =  "% latino",
    "XX_perc_white" = "% white",
    "XX_perc_asian" = "% asian",
    "XX_perc_island" = "% island native",
    "XX_perc_other" =   "% other",
    "XX_perc_two_more_races" = "% two more races",
    "XX_days_f0" =  "Number of days since first case"
  )
  )

tab_us_all <- inner_join(tab_us_tot, tab_us_flu, by = "var")

print(
  xtable(
    tab_us_all
  ),
  include.rownames=FALSE,
  comment = FALSE,
  align = rep("c", ncol(tab_it))
)

```

# Some figures US

```{r fig.width=12, fig.height=12}
library(ggcorrplot)
M <-   
  today_dat %>% 
  select_if(is.numeric) %>% 
  cor( use ="pairwise.complete.obs")
ggcorrplot(M,
           hc.order = TRUE
           # pch.cex = 30,
           # tl.cex = 30
)
```

```{r fig.width=4, fig.height=12, message=FALSE, warning=FALSE}
ordered_levels <-   
  today_dat %>% 
  group_by(state) %>% 
  mutate(mean_z = mean(ZZ_perc_imm65)) %>%
  arrange(mean_z) %>% 
  ungroup() %>%
  select(state) %>% 
  unique() %>% 
  unlist()

today_dat %>% 
  mutate(state = factor(state, levels = ordered_levels)) %>% 
  mutate(urban = factor(recode(XX_urban,
                        `0` = "rural",
                         `1` = "urban"
                        ))
         ) %>%

  ggplot(aes(ZZ_perc_imm65, state, fill = urban)) +
      geom_density_ridges2(alpha = 0.5, scale = 1.1) +
      labs(x = "influenza vaccination in 65 and older (percent)",
           y = NULL
           ) +
  theme(legend.title = element_blank(),
        legend.position="top")
```
```{r}
today_dat %>% 
  mutate(state = factor(state, levels = ordered_levels)) %>% 
  mutate(urban = factor(recode(XX_urban,
                        `0` = "rural",
                         `1` = "urban"
                        ))
         ) %>%

  ggplot(aes(ZZ_perc_imm65, fill = urban)) +
      geom_density(alpha = 0.5) +
      labs(x = "influenza vaccination in 65 and older (percent)",
           y = NULL
           ) +
  theme(legend.title = element_blank(),
        legend.position="top")
```

```{r fig.width=4, fig.height=12, message=FALSE, warning=FALSE}
ordered_levels <-   
  today_dat %>% 
  group_by(state) %>% 
  mutate(mean_z = mean(YY_deaths)) %>%
  arrange(mean_z) %>% 
  ungroup() %>%
  select(state) %>% 
  unique() %>% 
  unlist()
today_dat %>% 
  mutate(state = factor(state, levels = ordered_levels)) %>% 
  mutate(state = factor(state, levels = ordered_levels)) %>% 
  mutate(urban = factor(recode(XX_urban,
                        `0` = "rural",
                        `1` = "urban"
                        ))
         ) %>% 
  ggplot(aes(log(YY_deaths), state, fill = urban)) +
      geom_density_ridges2(alpha = 0.5) +
      labs(x = "number of deaths (log2)") +
     theme(legend.title = element_blank(),
           legend.position="top")

linbrary(G)
```

```{r}
today_dat %>% 
   group_by(state) %>% 
   summarize(deaths  = sum(YY_deaths)) %>%

   ggplot(aes(reorder(state, deaths), deaths)) +
     geom_col() +
     scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
     labs(y = NULL)

```

```{r}
today_dat %>% 
   group_by(date) %>% 
   mutate(tot = sum(YY_deaths)) %>% 
   group_by(state) %>%
   summarize(perc_death = round((sum(YY_deaths)/first(tot) * 100), 2)) %>% 
   arrange(desc(perc_death))
```

```{r}
today_dat %>% 
   group_by(date) %>% 
   mutate(tot = sum(NC_cases)) %>% 
   group_by(state) %>%
   summarize(perc_cases = round((sum(NC_cases)/first(tot) * 100),2)) %>% 
   arrange(desc(perc_cases))
```

```{r}
options(scipen = 999)
today_dat %>% 
   group_by(state) %>% 
   summarize(cases  = sum(NC_cases)) %>%

   ggplot(aes(reorder(state, cases), cases)) +
     geom_col() +
     scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
    labs(y = NULL)

```

```{r fig.width=4, fig.height=9}
today_dat %>% 
  group_by(state) %>% 
  summarize(temp_winter = mean(XX_winter_hum, na.rm = TRUE),
            n = n()
            ) %>%

  ggplot(aes(reorder(state, temp_winter), temp_winter)) +
    geom_col() +
   # scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 90)) +
   geom_label(aes(label = n,  y = 25)) +
   coord_flip()

```

```{r fig.width=4, fig.height=9}
today_dat %>% 
  group_by(state) %>% 
  summarize(temp_summer = mean(XX_summer_hum, na.rm = TRUE),
            n = n()
            ) %>%

  ggplot(aes(reorder(state, temp_summer), temp_summer)) +
    geom_col() +
   # scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 90)) +
   geom_label(aes(label = n,  y = 25)) +
   coord_flip()

```

```{r}
today_dat %>% 
  group_by(state) %>% 
  summarize(temp_summer = mean(XX_summer_hum, na.rm = TRUE),
            n = n()) %>%

  ggplot(aes(reorder(state, temp_summer), temp_summer)) +
    geom_col() +
   scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 90))

```

```{r}
today_dat %>% 
  filter(state == "Oregon") %>% 
  select(fips, XX_summer_hum) %>% 
  dput()
```

```{r}
today_dat %>% 
  group_by(state) %>% 
  rename(temp_winter = XX_winter_hum) %>%

  ggplot(aes(reorder(state, temp_winter), temp_winter, color = state)) +
    geom_jitter() +
   scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 90)) +
   theme(legend.position = "none")

```

```{r}
library(vroom)
us_season <- vroom("https://raw.githubusercontent.com/wxwx1993/PM_COVID/master/Data/temp_seasonal_county.csv",
  col_types = cols(.default = "d")
)

summary(us_season$winter_rmax)

```