### 2020-07-01 notebooks/old_notebooks/logo.Rmd

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}

library(devtools)

# devtools::install_github("c1au6i0/covid19census")
library(kableExtra)
library(covid19census)

ff <- getus_all()

ff <- ff[1:50, 1:50]

names(ff) <- paste0("x", 1:50)

```

```{r}
library(dplyr)
ff %>% 
  kable(escape = F, align = "c") %>% 
  column_spec(1:50, bold = T, border_right = T) %>%
  kable_styling(latex_options = c("striped", "scale_down"), full_width = FALSE, font_size = 5)

```

```{r}
library(dplyr)
it <- getit_all()
glimpse(it)
```

```{r}
library(showtext)
library(hexSticker)
library(ggplot2)
library(png)
library(grid)
imgurl <- readPNG("/Users/heverz/Desktop/logo.png")
imgurl <- rasterGrob(imgurl, interpolate=TRUE)

font_add_google("Gloria Hallelujah", "gloria")
showtext_auto()

p <- sticker(imgurl,
          package="covid19census", p_size=5, p_y = 1.5, s_x = 1.1, s_y = 0.7, s_width=2.3, s_height=2.3, p_color = "#FF6A6A", p_family = "gochi",
          filename="hexsticker.png", h_color = "#FF6A6A", h_fill="white")
p
```